---
title: "Markdown Template"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: haddock
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, warning=F, fig.width = 10)
```


# Load Packages

```{r}

library( haven )
library( tidyr )
library( dplyr )
library( ggplot2 )
library( ggthemes )
library( scales )
library( stargazer )
library( pander )

```

## Custom Functions

```{r}

jplot <- function( x1, x2, lab1="", lab2="", draw.line=T, ... )
{

	plot( x1, x2,
	      pch=19, 
	      col=gray(0.6, alpha = 0.5), 
	      cex=1.5,  
	      bty = "n",
	      xlab=lab1, 
	      ylab=lab2, cex.lab=1.5,
        ... )

	if( draw.line==T ){ 
		ok <- is.finite(x1) & is.finite(x2)
		lines( lowess(x2[ok]~x1[ok]), col="red", lwd=3 ) }

}

```



# Herfindahl-Hirschman Index

## Definitions

Herfindahl-Hirschman Index (HHI) is one of the most popular measures of market structure in nonprofit research and economics. Specifically, it is a measure of market concentration that is used to understand market competitiveness and organizations¡¦ competitive behavior and performance in markets.

## Calculations

HHI can be calculated as follows: 

Approach 1:

The index could be calculated as the sum of the squared market share of each firm competing in a market (Thornton & Belski, 2010). In the nonprofit sector, market share is determined by the ratio of a nonprofit¡¦s total revenue to the aggregate revenue for the organization's market. The calculated values range from 1/N to 1, where N is the number of organizations in a market. A market with an HHI of 1 is considered as a monopoly. In contrast, a market with an HHI close to 0 would be viewed as a nearly competition one.

Approach 2:

Another approach to calculate HHI used the same definition of market share as Approach 1; however, the market share of each organization is expressed as a whole number rather then a decimal in the calculation (Seaman, Wilsker, & Young, 2014; Soule & King, 2008; Thornoton, 2006; Castaneda, Garen, & Thornton, 2008). For example, we take the value of 40 to calculate HHI if a nonprofit¡¦s market share is 0.4 (40%). The calculated values range from 1/N to 10,000. A market with an HHI of 10,000 is considered as a monopoly. In contrast, a market with an HHI close to 0 would be viewed as a nearly competition one. The U.S. Department of Justice suggests that a market with an HHI less than 1,500 to be an unconcentrated marketplace, an HHI of 1,500 to 2,500 to be a moderately concentrated marketplace, and an HHI greater than 2,500 to be a highly concentrated marketplace. There is only scale difference Approach 1 and 2.

Approach 3:

Normalized HHI ranges between 0 to 1, rather than 1/N to 1 or 1/N to 10,000.

## Notes

We define a market as a Metropolitan Statistical Area (MSA) in which relevant nonprofits compete for resources to survive. The code to create HHI metrics is as follows.


# Create Metrics

```{r, eval=F}


# dat <- read.csv( "https://www.dropbox.com/s/xjad2z6zewwemq9/nccs.core2010pc.csv?dl=1" )

# setwd("C:/Users/JK/Desktop/R Practice")

core.2000 <- readRDS( "nccs.core2010pc.rds" )

names( core.2000 ) <- toupper( names( core.2000 ))

core <- select( core.2000, EIN, MSA_NECH, NTMAJ12, TOTREV)

core %>% group_by( MSA_NECH, NTMAJ12 ) %>% summarize( hhi= sum( (TOTREV / sum(TOTREV))^2 ) )

core %>% group_by( MSA_NECH, NTMAJ12 ) %>% summarize( Nhhi= sum((((TOTREV / sum(TOTREV))^2) - (1 / EIN)) / (1 -(1 / EIN))))


## Descriptive Statistics

### Average HHI in MSAs

core %>%
  group_by(MSA_NECH) %>%
  summarize( hhi= sum( (TOTREV / sum(TOTREV))^2 ) ) 


### Average HHI in Sub-Sectors

core %>%
  group_by(NTMAJ12) %>%
  summarize( hhi= sum( (TOTREV / sum(TOTREV))^2 ) )


head(core) %>% pander()

```


## Descriptive Statistics

```{r}

library( ggplot2 )


hhi.msa <- 
core %>% 
  group_by( MSA_NECH ) %>% 
  summarize( hhi= sum( (TOTREV / sum(TOTREV))^2 ) ) 
                       
qplot( data=hhi.msa, x=hhi, geom="histogram", xlim=c(0,1) ) 




nhhi <- 
core %>% group_by( MSA_NECH, NTMAJ12 ) %>% summarize( Nhhi= sum((((TOTREV / sum(TOTREV))^2) - (1 / EIN)) / (1 -(1 / EIN))))


qplot( data=nhhi, x=Nhhi, geom="histogram", xlim=c(0,1) ) + facet_wrap( ~ NTMAJ12, ncol=4 )





```



```{r}


ggplot( data=nhhi, aes(Nhhi)) + xlim( 0.0, 1.0 ) + geom_histogram()

ggplot( data=nhhi, aes( x=value, fill=Nhhi )) + 
        geom_density(alpha = 0.5) + xlim( 0.0, 1.0 ) +
        xlab( "Vital Rates" )  + theme_minimal()

ggplot( nhhi, aes( x=value, fill=Nhhi )) + 
        geom_density(alpha = 0.5) + xlim( 0.0, 1.0 ) +
        xlab( "Vital Rates" ) + facet_grid( variable ~ rate  ) + theme_minimal()

```



## Histograph Example

```{r, eval=F}

ggplot( df.vitals, aes( x=value, fill=rate )) + 
        geom_density(alpha = 0.5) + xlim( -0.10, 0.25 ) +
        xlab( "Vital Rates" ) + facet_grid( variable ~ rate  ) + theme_minimal()




new_theme <- function( base_size = 11, base_family = "" ) 
{
    theme_bw( base_size = base_size, base_family = base_family) %+replace% 
             theme(  axis.ticks = element_blank(), legend.background = element_blank(), 
             legend.key = element_blank(), panel.background = element_blank(),
             legend.position="none",
             axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank(),
             panel.border = element_blank(), strip.background = element_blank(), 
             plot.background = element_blank(), complete = TRUE )
}


ggplot( df.vitals, aes( x=value, fill=rate )) + 
        geom_density(alpha = 0.5) + xlim( -0.10, 0.25 ) +
        xlab( "Vital Rates" ) + 
        geom_segment( aes(x=0, y=0, xend=0, yend=5 ), color="gray20", lwd=1 ) +
        facet_grid( variable ~ rate  ) + 
        ggtitle("Vital Rates Across Metro Areas") + new_theme()

```





# Reference



Seaman, B., A. Wilsker, and D. R. Young. 2014. ¡¥¡¥Measuring Concentration and Competition in the U.S. Nonprofit Sector: Implications for Research and Public Policy.¡¦¡¦ Nonprofit Policy Forum 5(2): 231¡V259.

Thornoton, J. 2006. ¡¥¡¥Nonprofit Fund-Raising in Competitive Donor Markets.¡¦¡¦ Nonprofit and Voluntary Sector Quarterly 35: 204¡V224.

Soule, S. A., & King, B. G. (2008). Competition and resource partitioning in three social movement industries. American Journal of Sociology, 113(6), 1568-1610.

Castaneda, M. A., Garen, J., & Thornton, J. (2007). Competition, contractibility, and the market for donors to nonprofits. The Journal of Law, Economics, & Organization, 24(1), 215-246.

Thornton, J. P., & Belski, W. H. (2010). Financial reporting quality and price competition among nonprofit firms. Applied economics, 42(21), 2699-2713.







```{css, echo=F}
p {
color: black;
margin: 0 0 20px 0;
}

td {
    padding: 3px 10px 3px 10px;
    text-align: center;
}

table
{ 
    margin-left: auto;
    margin-right: auto;
    margin-top:80px;
    margin-bottom:100px;
}

h1, h2{
  margin-top:100px;
  margin-bottom:20px;
}

H5{
    text-align: center;
    color: gray;
    font-size:0.8em;
}

img {
    max-width: 90%;
    display: block;
    margin-right: auto;
    margin-left: auto;
    margin-top:30px;
    margin-bottom:20px;
}

pre {
  overflow-x: auto;
}

pre code {
   display: block; 
   padding: 0.5em;
   margin-bottom:20px;
}

code {
  font-size: 92%;
  border: 10px solid #F8F8F8;
  margin-bottom: 2px;
}

code[class] {
  background-color: #F8F8F8;
}

```

