---
title: "Markdown Template"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: haddock
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, warning=F, fig.width = 10)
```


# Load Packages

```{r}

library( haven )
library( tidyr )
library( dplyr )
library( ggplot2 )
library( ggthemes )
library( scales )
library( stargazer )
library( pander )
library( censusapi )


```

## Custom Functions

```{r}

jplot <- function( x1, x2, draw.line=T, ... )
{

	plot( x1, x2,
	      pch=19, 
	      col=gray(0.6, alpha = 0.4), 
	      cex=0.8,  
	      bty = "n",
	      cex.lab=1.5,
        ... )

	if( draw.line==T ){ 
		ok <- is.finite(x1) & is.finite(x2)
		lines( lowess(x2[ok]~x1[ok]), col="red", lwd=3 ) }

}

```



# H



## Definitions



## Calculations

can be calculated as follows: 

Approach 1:


Approach 2:



Approach 3:



## Notes

We define a market as a Metropolitan Statistical Area (MSA) in which relevant nonprofits compete for resources to survive. The code to create HHI metrics is as follows.


# API Census Data

```{r, eval=F}


# devtools::install_github("hrecht/censusapi")

censuskey <- "378079ddfd58f4b93006bb559f6eeda4d1bb15dc"

acs5_2010 <- getCensus( name="acs5", 
                        vintage=2010, 
                        key="378079ddfd58f4b93006bb559f6eeda4d1bb15dc", 
                        vars=c("B01001_001E","B19013_001E","B19301_001E",
                               "B01001A_001E","B01001B_001E",
                               "B01001D_001E", "B01001I_001E", "B01001_002E", 
                               "B01001_026E","B01002A_001E" ),
                               # ,"B23025_002E", 
                               #"B23025_004E","B23025_005E","B17010_017E","B17001_002E",
                               #"B09001_001E","B99211_002E" ), 
                        region="county:*")

head( acs5_2010 )


acs5_2010 <- acs5_2010 %>% 
          rename(Total Population = B01001_001E,  )




```

# Descriptive Statistics

## by MSAs

```{r}





```

##  by Subsector

```{r}




```



## and Subsector Size

```{r}








```

## and Total Revenue


```{r}




```


## HHI and Subsector Contribution


```{r}


```



# Reference











```{css, echo=F}
p {
color: black;
margin: 0 0 20px 0;
}

td {
    padding: 3px 10px 3px 10px;
    text-align: center;
}

table
{ 
    margin-left: auto;
    margin-right: auto;
    margin-top:80px;
    margin-bottom:100px;
}

h1, h2{
  margin-top:100px;
  margin-bottom:20px;
}

H5{
    text-align: center;
    color: gray;
    font-size:0.8em;
}

img {
    max-width: 90%;
    display: block;
    margin-right: auto;
    margin-left: auto;
    margin-top:30px;
    margin-bottom:20px;
}

pre {
  overflow-x: auto;
}

pre code {
   display: block; 
   padding: 0.5em;
   margin-bottom:20px;
}

code {
  font-size: 92%;
  border: 10px solid #F8F8F8;
  margin-bottom: 2px;
}

code[class] {
  background-color: #F8F8F8;
}

```

