---
title: "Metric 01 - HHI"
output:
  html_document:
    theme: readable
    df_print: paged
    highlight: tango
    toc: yes
    toc_float: no
    css: html/simple.css
    includes:
      after_body: html/footer.html
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, warning=F, fig.width = 10)
```







# Herfindahl-Hirschman Index



## Definitions

Herfindahl-Hirschman Index (HHI) is one of the most popular measures of market structure in nonprofit research and economics. Specifically, it is a measure of market concentration that is used to understand market competitiveness and organizations?? competitive behavior and performance in markets.

## Metric Calculations

HHI can be calculated as follows: 

**Approach 1:**

The index could be calculated as the sum of the squared market share of each firm competing in a market (Thornton & Belski, 2010). In the nonprofit sector, market share is determined by the ratio of a nonprofit??s total revenue to the aggregate revenue for the organization's market. The calculated values range from 1/N to 1, where N is the number of organizations in a market. A market with an HHI of 1 is considered as a monopoly. In contrast, a market with an HHI close to 0 would be viewed as a nearly competition one.

$$H =\sum_{i=1}^N s_i^2$$

*where $s_i$ is the market share of firm i in the market, and N is the number of firms.* 

*Thus, in a market with two firms that each have 50 percent market share, the Herfindahl index equals 0.502+0.502 = 1/2.* [wikipedia](https://en.wikipedia.org/wiki/Herfindahl%E2%80%93Hirschman_Index)


**Approach 2:**

Another approach to calculate HHI used the same definition of market share as Approach 1; however, the market share of each organization is expressed as a whole number rather then a decimal in the calculation (Seaman, Wilsker, & Young, 2014; Soule & King, 2008; Thornoton, 2006; Castaneda, Garen, & Thornton, 2008). For example, we take the value of 40 to calculate HHI if a nonprofit??s market share is 0.4 (40%). The calculated values range from 1/N to 10,000. A market with an HHI of 10,000 is considered as a monopoly. In contrast, a market with an HHI close to 0 would be viewed as a nearly competition one. The U.S. Department of Justice suggests that a market with an HHI less than 1,500 to be an unconcentrated marketplace, an HHI of 1,500 to 2,500 to be a moderately concentrated marketplace, and an HHI greater than 2,500 to be a highly concentrated marketplace. There is only scale difference Approach 1 and 2.

**Approach 3:**

Normalized HHI ranges between 0 to 1, rather than 1/N to 1 or 1/N to 10,000.

## Notes

We define a market as a Metropolitan Statistical Area (MSA) in which relevant nonprofits compete for resources to survive. The code to create HHI metrics is as follows.


# Packages

```{r}

library( haven )        # importing data files 
library( tidyr )        # data wrangling
library( dplyr )        # data wrangling 
library( ggplot2 )      # fancy plots 
library( ggthemes )     # fancy plots
library( scales )       # re-scaling numbers
library( stargazer )    # nice tables 
library( pander )       # format tables for HTML 
library( DT )           # embed datasets in HTML docs

source( "r-helper-functions/helper-functions.R" )
```


```{r, echo=F}
# stargazer table print type
# run chunk for live RMD sessions
s.type <- "text"
```


```{r, echo=F}
# stargazer table print type
# resets to html when knitting
s.type <- "html"
```

# Data

```{r, cache=TRUE}
core.2010 <- readRDS( "Data/nccs.core2010pc.rds" )

names( core.2010 ) <- toupper( names( core.2010 ))

core <- select( core.2010, EIN, MSA_NECH, NTMAJ12, TOTREV, CONT )

core$NTMAJ12[ core$NTMAJ12 == "AR" ] <- "Arts"
core$NTMAJ12[ core$NTMAJ12 == "BH" ] <- "Universities"
core$NTMAJ12[ core$NTMAJ12 == "ED" ] <- "Education"
core$NTMAJ12[ core$NTMAJ12 == "EH" ] <- "Hospitals"
core$NTMAJ12[ core$NTMAJ12 == "EN" ] <- "Environmental"
core$NTMAJ12[ core$NTMAJ12 == "HE" ] <- "Health"
core$NTMAJ12[ core$NTMAJ12 == "HU" ] <- "Human Services"
core$NTMAJ12[ core$NTMAJ12 == "IN" ] <- "International"
core$NTMAJ12[ core$NTMAJ12 == "MU" ] <- "Mutual Benefit"
core$NTMAJ12[ core$NTMAJ12 == "PU" ] <- "Public Benefit"
core$NTMAJ12[ core$NTMAJ12 == "RE" ] <- "Religion"
core$NTMAJ12[ core$NTMAJ12 == "UN" ] <- "Unknown"

core$NTMAJ12 <- factor( core$NTMAJ12 )

head(core) %>% pander()
```


Recode negative revenues as zero because they cause HHIs above 1:

```{r}
core$TOTREV3 <- core$TOTREV
core$TOTREV[ core$TOTREV < 0 ] <- 0
```


```{r}
msa.pop <- read.csv("Data/msa-pop.csv")
```



# Create Competition Metrics


## Regular HHI

```{r, eval=T}
dat.hhi <- 
  core %>% 
  group_by( MSA_NECH, NTMAJ12 ) %>% 
  summarize( hhi= sum( (TOTREV / sum(TOTREV))^2 ), 
             n=n(), 
             revenue= sum(TOTREV), 
             contribution=sum(CONT) )

dat.hhi$hhi[ dat.hhi$hhi > 1 ] <- 1

dat.hhi$revenue[ dat.hhi$revenue < 0 ] <- 0

dat.hhi <- merge( dat.hhi, msa.pop, all.x=T )
```


```{r, results="asis"}
dat.hhi %>%
  # rescale for printing
  mutate( hhi = 100 * hhi,
          revenue_thousands = revenue / 1000, 
          contributions_thousands = contribution / 1000 ) %>%  
  select( hhi, n, revenue_thousands, contributions_thousands, MSA_POP ) %>%
stargazer( type = s.type, 
           digits=0, 
           summary.stat = c("min","p25","median","mean","p75","max", "sd"))
```




## Normalized HHI

```{r}
# There is a clarification called TOTREV2 on the list. 
# What is the difference between TOTREV and TOTREV2?  
# CONT represents total contributions.


dat.nhhi <-
  core %>% 
  group_by( MSA_NECH, NTMAJ12 ) %>% 
  summarize( 
    nhhi= sum((((TOTREV / sum(TOTREV))^2) - 
                 (1 / EIN)) / (1 -(1 / EIN))), 
    n=n(), revenue= sum(TOTREV))


dat.nhhi$nhhi[ dat.nhhi$nhhi > 1 ] <- 1

dat.nhhi$revenue[ dat.nhhi$revenue < 0 ] <- 0
```

```{r, results="asis"}
stargazer( as.data.frame(dat.nhhi), type=s.type )
```

```{r}
dat.nhhi %>%
  group_by( NTMAJ12 ) %>%
  summarize( n=n(), min=min(nhhi), max=max(nhhi) ) %>% 
  pander()
```




### Data Download 

```{r, fig.height=10}
# library( DT )

these.buttons <- c( 'copy', 'csv', 'excel', 'print' )

datatable( dat.nhhi,
           filter='bottom', rownames=FALSE, 
           #options=list( pageLength=5, autoWidth=TRUE ),
           fillContainer=TRUE, 
           style="bootstrap",
           class='table-condensed table-striped',
           extensions = 'Buttons', 
           options=list( dom='Bfrtip', 
                         buttons=these.buttons  )) %>%
  
formatStyle( "MSA_NECH", "white-space"="nowrap" )
```



# Descriptive Statistics


## HHI by MSAs

```{r}
dat.msa <-
  core %>%
  group_by(MSA_NECH) %>%
  summarize( hhi= sum( ( TOTREV / sum(TOTREV) )^2 ), 
             n=n(), 
             revenue= sum(TOTREV) ) 

# dat.msa$hhi[ dat.msa$hhi > 1 ] <- 1

# dat.msa$revenue[ dat.msa$revenue < 0 ] <- 0

summary( dat.msa ) %>% pander()


ggplot( dat.msa, aes(x = hhi )) +  
  geom_density( alpha = 0.5, fill="blue" ) + 
  xlim( 0, 1 ) 
```



## HHI by MSA Population

```{r}
# msa.pop <- read.csv("Data/MSA_POP.csv")
# msa.names <- read.csv("Data/msa-names.csv")
# msa.pop <- merge( msa.pop, msa.names, all.x=TRUE )

dat.msa <- merge( dat.msa, msa.pop, by="MSA_NECH", all.x=T )

jplot( log10(dat.msa$MSA_POP), 
       dat.msa$hhi, xaxt="n",
       xlab="MSA Population", ylab="HHI"  )
axis( side=1, 
      at=c(0,1,2,3,4), 
      labels=c("1","10","100","1,000","10,000") )



ggplot( dat.msa, aes(log10(MSA_POP),hhi) )  + 
        geom_point( col="gray30", alpha=0.7) + 
        xlab( "MSA Population (logged)" ) + 
        ylab( "HHI Across MSAs" ) +
        theme_minimal()
```


## HHI by Subsector

```{r}

dat.sub <-
  core %>%
  group_by(NTMAJ12) %>%
  summarize( hhi= round( sum( (TOTREV / sum(TOTREV))^2 ), 3 ), 
             revenue= sum(TOTREV) ) %>%
  arrange( - hhi )  %>% pander( )

summary(dat.sub)

ggplot( dat.hhi, aes(x = hhi)) +  
  geom_density( alpha = 0.5, fill="blue" ) + 
  xlim( -0.05, 1 ) +
  xlab( "HHI" )


ggplot( dat.hhi, aes( x=hhi ) ) + 
  geom_density( alpha = 0.5, fill="blue" ) + 
  xlim( -0.05, 1 ) +
  xlab( "HHI" ) + facet_wrap( ~ NTMAJ12, nrow=3 ) + 
  theme_minimal()

ggplot( dat.hhi, aes(y = hhi) )  + 
  geom_boxplot( col="gray30", alpha=0.7) + 
  ylab( "HHI Across MSAs" ) +
  facet_wrap( ~ NTMAJ12, nrow=3 ) + 
  theme_minimal()

```




## HHI and Subsector Size

```{r}

jplot( log10(dat.hhi$n), dat.hhi$hhi, xaxt="n",
       xlab="Number of Nonprofits", ylab="HHI",  )
axis( side=1, 
      at=c(0,1,2,3,4), 
      labels=c("1","10","100","1,000","10,000") )



ggplot( dat.hhi, aes(log10(n),hhi) )  + 
  geom_point( col="gray30", alpha=0.7) + 
  xlab( "Number of Nonprofits (logged)" ) + 
  ylab( "HHI Across MSAs" ) +
  theme_minimal()


ggplot( dat.hhi, aes(log10(n),hhi) )  + 
  geom_point( col="gray30", alpha=0.7) + 
  xlab( "Number of Nonprofits (logged)" ) + 
  ylab( "HHI Across MSAs" ) +
  facet_wrap( ~ NTMAJ12, nrow=3 ) + 
  theme_minimal()

```

## HHI and Total Revenue


```{r}

jplot( log10(dat.hhi$revenue), dat.hhi$hhi, xaxt="n",
       xlab="Total Revenue (logged)", ylab="HHI"  )
axis( side=1, 
      at=c(0,1,2,3,4,5,6,7,8,9,10,11), 
      labels=c("1","10","100","1K","10K","100K",
               "1M","10M","100M","1B","10B","100B") )




ggplot( dat.hhi, aes(log10(revenue),hhi) )  + 
  geom_point( col="gray30", alpha=0.7) +
  xlab( "Total Revenue (logged)" ) + 
  ylab( "HHI Across MSAs" ) +
  theme_minimal()


ggplot( dat.hhi, aes(log10(revenue),hhi) )  + 
  geom_point( col="gray30", alpha=0.7) +
  xlab( "Total Revenue (logged)" ) + 
  ylab( "HHI Across MSAs" ) +
  facet_wrap( ~ NTMAJ12, nrow=3 ) + 
  theme_minimal()



ggplot( dat.hhi, aes(log10(revenue),hhi) )  + 
  geom_point( col="gray30", alpha=0.7) + 
  geom_smooth(method="lm") + 
  xlab( "Total Revenue (logged)" ) + ylab( "HHI Across MSAs" ) +
  facet_wrap( ~ NTMAJ12, nrow=3 ) + 
  theme_minimal()

```


## HHI and Subsector Contribution


```{r}

jplot( log10(dat.hhi$contribution), dat.hhi$hhi, 
       xlab="Contribution (logged)", ylab="HHI", 
       xaxt="n" ) 
axis( side=1, 
      at=c(0,1,2,3,4,5,6,7,8,9,10,11), 
      labels=c("1","10","100","1K","10K","100K",
               "1M","10M","100M","1B","10B","100B") )


ggplot( dat.hhi, aes(log10(contribution),hhi) )  + 
  geom_point( col="gray30", alpha=0.7) +
  xlab( "Contrubution (logged)" ) + 
  ylab( "HHI Across MSAs" ) +
  theme_minimal()


ggplot( dat.hhi, aes(log10(contribution),hhi) )  +
  geom_point( col="gray30", alpha=0.7) +
  xlab( "Contrubution (logged)" ) + 
  ylab( "HHI Across MSAs" ) +
  facet_wrap( ~ NTMAJ12, nrow=3 ) + 
  theme_minimal()


ggplot( dat.hhi, aes(log10(contribution),hhi) )  + 
  geom_point( col="gray30", alpha=0.7) +
  geom_smooth(method="lm") + 
  xlab( "Contrubution (logged)" ) + 
  ylab( "HHI Across MSAs" ) +
  facet_wrap( ~ NTMAJ12, nrow=3 ) + 
  theme_minimal()
```



# Reference



Seaman, B., A. Wilsker, and D. R. Young. 2014. ????Measuring Concentration and Competition in the U.S. Nonprofit Sector: Implications for Research and Public Policy.???? Nonprofit Policy Forum 5(2): 231?V259.

Thornoton, J. 2006. ????Nonprofit Fund-Raising in Competitive Donor Markets.???? Nonprofit and Voluntary Sector Quarterly 35: 204?V224.

Soule, S. A., & King, B. G. (2008). Competition and resource partitioning in three social movement industries. American Journal of Sociology, 113(6), 1568-1610.

Castaneda, M. A., Garen, J., & Thornton, J. (2007). Competition, contractibility, and the market for donors to nonprofits. The Journal of Law, Economics, & Organization, 24(1), 215-246.

Thornton, J. P., & Belski, W. H. (2010). Financial reporting quality and price competition among nonprofit firms. Applied economics, 42(21), 2699-2713.

Mendoza-Abarca, K. I., & Gras, D. (2017). The Performance Effects of Pursuing a Diversification Strategy by Newly Founded Nonprofit Organizations. Journal of Management, 0149206316685854.



<br>
<br>
<hr>
<br>
<br>




```{css, echo=F}

div#TOC{ margin-top: 42px; }
```

