---
title: "Markdown Template"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: haddock
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, warning=F, fig.width = 10)

```


# Load Packages

```{r}

library( haven )
library( tidyr )
library( dplyr )
library( ggplot2 )
library( ggthemes )
library( scales )
library( stargazer )
library( pander )

```

## Custom Functions

```{r,echo=F}

jplot <- function( x1, x2, draw.line=T, ... )
{

	plot( x1, x2,
	      pch=19, 
	      col=gray(0.6, alpha = 0.5), 
	      cex=1.5,  
	      bty = "n",
	      cex.lab=1.5,
        ... )

	if( draw.line==T ){ 
		ok <- is.finite(x1) & is.finite(x2)
		lines( lowess(x2[ok]~x1[ok]), col="red", lwd=3 ) }

}

```



# Fundraising Efficiency 


## Definitions

Fundraising efficiency is the measure of the fundraising competition in a market. It represents total private donations raised relative to expenses on fundraising activities and indicates the amount of private donations raised for each dollar of fundraising expense incurred. Lower ratio indicates greater fundraising competition in a market.  

## Calculations

Fundraising efficiency be calculated as follows: 

Private donations divided by fundraising expenses (Brooks, 2004; Greenlee & Bukovinsky, 1998; Ritchie & Kolodinsky, 2003)


## Notes

We define a market as a Metropolitan Statistical Area (MSA) in which relevant nonprofits compete for resources to survive. 




# Create Metrics

## E-File Exploratory

```{r, cache=T}

core.2010 <- readRDS(gzcon(url("https://www.dropbox.com/s/byt7kh3kejg4ycl/nccs.core2010pc.rds?dl=1")))

efile <- readRDS(gzcon(url("https://www.dropbox.com/s/2taibiudu0dh4j3/2014-EFILE-CORE.rds?dl=1")))

pop <- read.csv( "https://www.dropbox.com/s/k0b8o17k03ffio0/MSA.Population.2010.csv?dl=1", stringsAsFactors=F )


nrow( efile )

sum( efile$TOTFUNDEXP > 0, na.rm=T  )

cont <- as.numeric( efile$CONTRIBCURRENT )
expense <- as.numeric( efile$TOTFUNDEXP )

expense[ expense <= 0 ] <- NA
efficiency <- cont / expense 

efficiency[ efficiency > 500 ] <- NA

summary( efficiency )


# ADD MSA CODES FROM CORE 2010

crosswalk <- select( core.2010, MSA_NECH, EIN )

efile$EIN <- as.numeric( efile$EIN )

efile <- merge( efile, crosswalk, all.x=T )




# Efficiency by MSA

MSA <- efile$MSA_NECH
MSA[ as.numeric( efile$TOTFUNDEXP ) <= 0 ] <- NA

tapply( efficiency, MSA, median, na.rm=T ) %>% summary()
tapply( efficiency, MSA, mean, na.rm=T ) %>% summary()


t1 <- tapply( efficiency, MSA, median, na.rm=T ) %>% as.numeric()
t2 <- table( MSA ) %>% as.numeric()

t2.log <- log(t2+1)


jplot( x1=t2.log, x2=t1, xlab="Num Nonprofits (log)", ylab="FE" )

jplot( x1=t2.log, x2=t1, xlab="Num Nonprofits (log)", ylab="Fundraising Efficiency", 
       ylim=c(0,25), xlim=c(4,10) )


```



## SOI Exploratory

```{r, cache=T}

dat <- read.csv( "https://nccs-data.urban.org/data/soi/2012/soi.soi12c39.csv" )

summary( dat$tot_fndrsng_exp_cy )

sum( dat$tot_fndrsng_exp_cy > 0 )

nrow( dat )



fund <- dat$tot_fndrsng_exp_cy
cont <- dat$contri_grnts_cy

fund[ fund <= 0 ] <- NA

efficiency <- cont / fund

efficiency[ efficiency > 500 ] <- NA

summary( efficiency )

hist( efficiency, col="gray", border=NA, breaks=1000, xlim=c(0,50) )



# Efficiency by MSA

tapply( efficiency, dat$MSA_NECH, median, na.rm=T ) %>% summary()

t1 <- tapply( efficiency, dat$MSA_NECH, median, na.rm=T ) %>% as.numeric()
t2 <- table( dat$MSA_NECH ) %>% as.numeric()

t2.log <- log(t2+1)


jplot( x1=t2.log, x2=t1, xlab="Num Nonprofits (log)", ylab="FE" )

jplot( x1=t2.log, x2=t1, xlab="Num Nonprofits (log)", ylab="FE", ylim=c(0,25) )


```



## By MSAs


```{r, eval=T, cache=T}

# setwd( "C:/Users/JK/Desktop/R Practice" )

# core.2010 <- readRDS( "nccs.core2010pc.rds" )

# core.2010 <- readRDS(gzcon(url("https://www.dropbox.com/s/byt7kh3kejg4ycl/nccs.core2010pc.rds?dl=1")))



names( core.2010 ) <- toupper( names( core.2010 ))

core <- select( core.2010, MSA_NECH, NTMAJ12, CONT, SOLICIT, FUNDFEES, TOTREV )

# View(core)

summary(core)


dat.msa <-
core %>% 
  group_by( MSA_NECH ) %>% 
  summarize( SOL=sum(SOLICIT), cont=sum(CONT), FE = ( sum(CONT) / sum(SOLICIT) ), n=n() ) 

dat.msa$revenue[ dat.msa$revenue < 0 ] <- 0


dat.fe <- 
core %>% 
  group_by( MSA_NECH ) %>% 
  summarize( sol=sum(SOLICIT), 
             fundfees=sum(FUNDFEES),
             cont=sum(CONT),
             FE = ( sum(CONT, na.rm=T) / sum( FUNDFEES + SOLICIT, na.rm=T ) ), 
             n=n() ) 

dat.fe$cont[ dat.fe$cont < 0 ] <- 0
dat.fe$sol[ dat.fe$sol < 0 ] <- 0
dat.fe$fundfees[ dat.fe$fundfees < 0 ] <- 0



                                                      
ggplot( dat.fe, aes( x=FE ) ) + 
        geom_density( alpha = 0.5, fill="blue" ) + xlim( -1000, 10000 ) +
        xlab( "Fundraising Efficiency" ) + theme_minimal()

jplot <- function( x1, x2, draw.line=T, ... )
{

	plot( x1, x2,
	      pch=19, 
	      col=gray(0.6, alpha = 0.5), 
	      cex=1.5,  
	      bty = "n",
	      cex.lab=1.5,
        ... )

	if( draw.line==T ){ 
		ok <- is.finite(x1) & is.finite(x2)
		lines( lowess(x2[ok]~x1[ok]), col="red", lwd=3 ) }
  
}


jplot( log10(dat.fe$n), dat.fe$FE, xaxt="n", xlim=c(2,5), ylim=c(-100,9000),
       xlab="Number of Nonprofits", ylab="Fundraising Efficiency"  )
axis( side=1, at=c(2,3,4,5), labels=c("100","1,000","10,000","100,000") )


```


## By Subsector

```{r}


core$NTMAJ12[ core$NTMAJ12 == "AR" ] <- "Arts"
core$NTMAJ12[ core$NTMAJ12 == "BH" ] <- "Universities"
core$NTMAJ12[ core$NTMAJ12 == "ED" ] <- "Education"
core$NTMAJ12[ core$NTMAJ12 == "EH" ] <- "Hospitals"
core$NTMAJ12[ core$NTMAJ12 == "EN" ] <- "Environmental"
core$NTMAJ12[ core$NTMAJ12 == "HE" ] <- "Health"
core$NTMAJ12[ core$NTMAJ12 == "HU" ] <- "Human Services"
core$NTMAJ12[ core$NTMAJ12 == "IN" ] <- "International"
core$NTMAJ12[ core$NTMAJ12 == "MU" ] <- "Mutual Benefit"
core$NTMAJ12[ core$NTMAJ12 == "PU" ] <- "Public Benefit"
core$NTMAJ12[ core$NTMAJ12 == "RE" ] <- "Religion"
core$NTMAJ12[ core$NTMAJ12 == "UN" ] <- "Unknown"

core$NTMAJ12 <- factor( core$NTMAJ12 )
core$MSA_NECH <- factor( core$MSA_NECH )

core$CONT[ core$CONT < 0 ] <- NA
core$TOTFEES[ core$TOTFEES < 0 ] <- NA

core$TOTFEES <- core$FUNDFEES + core$SOLICIT
core$TOTFEES[ core$TOTFEES == 0 ] <- NA



# sum values by metro/subsector then calculate efficiency
# this is wrong!
# will over-emphasize outlier org data

dat.fe <- 
core %>% 
  group_by( MSA_NECH) %>% 
  summarize( sol=sum(SOLICIT), 
             cont=sum(CONT),
             fundfees=sum(FUNDFEES),
             totfees=sum(TOTFEES, na.rm=T ),
             FE = ( sum(CONT, na.rm=T) / sum( TOTFEES, na.rm=T ) ), 
             n=n() ) 



# calcualate efficiency by orgs
# then average over metro / subsector
# this is better because we want to know
# the average cose of raising money
# by city / subsector


core$FE <- core$CONT / core$TOTFEES

summary( core$TOTFEES )
summary( core$CONT )
summary( core$FE )



dat.ave <- 
  core %>% 
  group_by( MSA_NECH ) %>% 
  summarize( mean.fe = mean( FE, na.rm=T ), median.fe = median(FE, na.rm=T ), 
             max.fe = max( FE, na.rm=T ), n=n() )

ggplot( dat.ave, aes( x=median.fe ) ) + geom_density( alpha = 0.5, fill="blue" ) + xlim( -50, 300 )



ggplot( core, aes( x=FE ) ) + 
        geom_density( alpha = 0.5, fill="blue" ) + xlim( -50, 300 ) +
        xlab( "Fundraising Efficiency" ) + facet_wrap( ~ NTMAJ12, nrow=3 ) + theme_minimal()


View(core)


dat.ave3 <- 
  core %>% 
  group_by( MSA_NECH, NTMAJ12 ) %>% 
  summarize( mean.fe = mean( FE, na.rm=T ), median.fe = median(FE, na.rm=T ) )

ggplot( dat.ave3, aes( x=median.fe ) ) + 
        geom_density( alpha = 0.5, fill="blue" ) + xlim( -50, 300 ) +
        xlab( "Fundraising Efficiency" ) + facet_wrap( ~ NTMAJ12, nrow=3 ) + theme_minimal()


```



```{r, fig.height=40, fig.width=10, eval=F}

# ggplot( dat.fe, aes( x=FE ) ) + 
#         geom_density( alpha = 0.5, fill="blue" ) + xlim( -10, 50 ) +
#         xlab( "Fundraising Efficiency" ) + facet_wrap( ~ MSA_NECH, nrow=40 ) + theme_minimal()

ggplot( dat.fe, aes( x=NA, y=log10(FE) ) ) + 
        geom_boxplot( fill="gray30" ) + facet_wrap( ~ MSA_NECH, nrow=40 ) + theme_minimal()


```


## Fundraising Costs By NPO Size

```{r}

core <- select( core.2000, MSA_NECH, NTMAJ12, CONT, SOLICIT, FUNDFEES, TOTREV )

core$CONT[ core$CONT < 0 ] <- NA

core$TOTFEES <- core$FUNDFEES + core$SOLICIT
core$TOTFEES[ core$TOTFEES == 0 ] <- NA

core$FE <- core$CONT / core$TOTFEES




jplot <- function( x1, x2, draw.line=T, ... )
{

	plot( x1, x2,
	      pch=19, 
	      col=gray(0.6, alpha = 0.2), 
	      cex=0.8,  
	      bty = "n",
	      cex.lab=1.5,
        ... )

	if( draw.line==T ){ 
		ok <- is.finite(x1) & is.finite(x2)
		lines( lowess(x2[ok]~x1[ok]), col="red", lwd=3 ) }

}

jplot( log10(core$TOTREV), core$FE, ylim=c(0,500), xlim=c(4,10), xaxt="n",
       xlab="Nonprofit Revenue", ylab="Fundraising Efficiency")
axis( side=1, at=4:10, labels=c("$1k","$10k","$100k","$1m","$10m","$100m","$1b") )

```


## FE by nonprofit age

```{r}


core <- select( core.2000, MSA_NECH, NTMAJ12, CONT, SOLICIT, FUNDFEES, TOTREV, RULEDATE )

core$RULEYEAR <- (core$RULEDATE /100)
core$RULEYEAR <-round(core$RULEYEAR, 0)
core$RULEYEAR[ core$RULEYEAR == 0 ] <- NA
core$orgage <- 2011 - core$RULEYEA + 1
core$orgage[ core$orgage < 0 ] <- NA

summary(core$orgage)
View(core)
  
core$TOTFEES <- core$FUNDFEES + core$SOLICIT

View(core)

core$CONT[ core$CONT < 0 ] <- NA

core$TOTFEES <- core$FUNDFEES + core$SOLICIT
core$TOTFEES[ core$TOTFEES == 0 ] <- NA

core$FE <- core$CONT / core$TOTFEES




jplot <- function( x1, x2, draw.line=T, ... )
{

	plot( x1, x2,
	      pch=19, 
	      col=gray(0.6, alpha = 0.2), 
	      cex=0.8,  
	      bty = "n",
	      cex.lab=1.5,
        ... )

	if( draw.line==T ){ 
		ok <- is.finite(x1) & is.finite(x2)
		lines( lowess(x2[ok]~x1[ok]), col="red", lwd=3 ) }

}




jplot(core$orgage, core$FE, xlim=c(0,120), ylim=c(0,500), 
       xlab="Nonprofit Age", ylab="Fundraising Efficiency")




```


## Reference

Brooks, A. C. (2004). Evaluating the effectiveness of nonprofit fundraising. Policy Studies Journal, 32(3), 363-374.

Greenlee, J. S., & Bukovinsky, D. (1998). Financial ratios for use in the analytical review of charitable organizations. Ohio CPA Journal, 57, 32-38.

Ritchie, W. J., & Kolodinsky, R. W. (2003). Nonprofit organization financial performance measurement: An evaluation of new and existing financial performance measures. Nonprofit Management and Leadership, 13(4), 367-381.

Thornton, J. (2006). Nonprofit fund-raising in competitive donor markets. Nonprofit and Voluntary Sector Quarterly, 35(2), 204-224.





```{css, echo=F}
p {
color: black;
margin: 0 0 20px 0;
}

td {
    padding: 3px 10px 3px 10px;
    text-align: center;
}

table
{ 
    margin-left: auto;
    margin-right: auto;
    margin-top:80px;
    margin-bottom:100px;
}

h1, h2{
  margin-top:100px;
  margin-bottom:20px;
}

H5{
    text-align: center;
    color: gray;
    font-size:0.8em;
}

img {
    max-width: 90%;
    display: block;
    margin-right: auto;
    margin-left: auto;
    margin-top:30px;
    margin-bottom:20px;
}

pre {
  overflow-x: auto;
}

pre code {
   display: block; 
   padding: 0.5em;
   margin-bottom:20px;
}

code {
  font-size: 92%;
  border: 10px solid #F8F8F8;
  margin-bottom: 2px;
}

code[class] {
  background-color: #F8F8F8;
}

```

