---
title: "Markdown Template"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: haddock
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, warning=F, fig.width = 10)
```


# Load Packages

```{r}

library( haven )
library( tidyr )
library( dplyr )
library( ggplot2 )
library( ggthemes )
library( scales )
library( stargazer )
library( pander )


```

## Custom Functions

```{r}

jplot <- function( x1, x2, draw.line=T, ... )
{

	plot( x1, x2,
	      pch=19, 
	      col=gray(0.6, alpha = 0.4), 
	      cex=0.8,  
	      bty = "n",
	      cex.lab=1.5,
        ... )

	if( draw.line==T ){ 
		ok <- is.finite(x1) & is.finite(x2)
		lines( lowess(x2[ok]~x1[ok]), col="red", lwd=3 ) }

}

```



# The Concentration of Commercial Nonprofitd Index



## Definitions



## Calculations



Approach 1:



Approach 2:



Approach 3:



## Notes

We define a market as a Metropolitan Statistical Area (MSA) in which relevant nonprofits compete for resources to survive. The code to create the metrics is as follows.


# Create Metrics

```{r, eval=T}



# setwd("C:/Users/JK/Desktop/R Practice")

# core.2010 <- readRDS( "nccs.core2010pc.rds" )

core.2010 <- readRDS(gzcon(url("https://www.dropbox.com/s/byt7kh3kejg4ycl/nccs.core2010pc.rds?dl=1")))

pop <- read.csv( "https://www.dropbox.com/s/k0b8o17k03ffio0/MSA.Population.2010.csv?dl=1", stringsAsFactors=F )


names( core.2010 ) <- toupper( names( core.2010 ))

core <- select( core.2010, EIN, MSA_NECH, NTMAJ12, TOTREV, PROGREV, DUES, INVINC, SAVINT, SECINC, NETRENT, OTHINVST, SALESECN, SALEOTHN, GRPROF)


######
# PROGREV Program Service Revenue
# DUES    Dues
# INVINC  Total Investment Income
# SAVINT  Interest Revenue
# SECINC  Dividends
# NETRENT Net Rental Income
# OTHINVST Other Investment Income
# SALESECN Net Gain/Loss: Sale of Securities
# SALEOTHN Net Gain/Loss: Sale of Other Assets
# GRPROF   Inventory Gross Profit

core$NTMAJ12[ core$NTMAJ12 == "AR" ] <- "Arts"
core$NTMAJ12[ core$NTMAJ12 == "BH" ] <- "Universities"
core$NTMAJ12[ core$NTMAJ12 == "ED" ] <- "Education"
core$NTMAJ12[ core$NTMAJ12 == "EH" ] <- "Hospitals"
core$NTMAJ12[ core$NTMAJ12 == "EN" ] <- "Environmental"
core$NTMAJ12[ core$NTMAJ12 == "HE" ] <- "Health"
core$NTMAJ12[ core$NTMAJ12 == "HU" ] <- "Human Services"
core$NTMAJ12[ core$NTMAJ12 == "IN" ] <- "International"
core$NTMAJ12[ core$NTMAJ12 == "MU" ] <- "Mutual Benefit"
core$NTMAJ12[ core$NTMAJ12 == "PU" ] <- "Public Benefit"
core$NTMAJ12[ core$NTMAJ12 == "RE" ] <- "Religion"
core$NTMAJ12[ core$NTMAJ12 == "UN" ] <- "Unknown"

head(core) %>% pander()
             

core$NTMAJ12 <- factor( core$NTMAJ12 )


core$commercialrevenues = core$PROGREV + core$DUES + core$INVINC + core$SAVINT + core$SECINC + core$NETRENT + core$OTHINVST + core$SALESECN + core$SALEOTHN + core$GRPROF

core$commercialpercent = core$commercialrevenues/core$TOTREV*100

dat.commercial <- 
  core %>% 
  group_by( MSA_NECH) %>% 
  summarize(n=n(), revenue = sum(TOTREV), n_commercial=sum(commercialpercent > 50), 
            concentration=n_commercial/n*100 )

# View(dat.commercial)



dat.commercial.sub <- 
  core %>% 
  group_by( MSA_NECH, NTMAJ12) %>% 
  summarize(n=n(), revenue = sum(TOTREV), n_commercial=sum(commercialpercent > 50), 
            concentration=n_commercial/n*100 )

# View(dat.commercial.sub)

```

# Descriptive Statistics

## Big nonprofits by MSAs and MSAs/Sub

```{r}

dat.commercial <- 
  core %>% 
  group_by( MSA_NECH) %>% 
  summarize(n=n(), revenue = sum(TOTREV), n_commercial=sum(commercialpercent > 50), 
            concentration=n_commercial/n*100 )

ggplot( dat.commercial, aes(x=concentration)) +  geom_histogram()

ggplot( dat.commercial, aes(x=n_commercial)) +  geom_bar()

ggplot(dat.commercial, aes(x=concentration)) +  geom_density( alpha = 0.5, fill="blue" ) + xlim( 0, 100 )


dat.msa.sub <-
  core %>%
  group_by(MSA_NECH,NTMAJ12) %>%
  summarize(n=n(), revenue = sum(TOTREV), n_commercial=sum(commercialpercent > 50), 
            concentration=n_commercial/n*100 )

ggplot( dat.msa.sub, aes(x=concentration)) +  geom_histogram()

ggplot( dat.msa.sub, aes(x=n_commercial)) +  geom_bar() + xlim( 0, 1000 )


ggplot(dat.msa.sub, aes(x=concentration)) +  geom_density( alpha = 0.5, fill="blue" ) + xlim( 0, 100 )


```

## Commercial organization by Subsector

```{r}

dat.sub <-
          core %>%
          group_by(NTMAJ12) %>%
          summarize(n=n(), revenue = sum(TOTREV), n_commercial=sum(commercialpercent > 50), 
            concentration=n_commercial/n*100 ) 

summary(core$TOTREV)

dat.commercial.sub <- 
  core %>% 
  group_by( MSA_NECH, NTMAJ12) %>% 
  summarize(n=n(), revenue = sum(TOTREV), n_commercial=sum(commercialpercent > 50), 
            concentration=n_commercial/n*100 )

ggplot( dat.commercial.sub, aes(x = concentration)) +  
        geom_density( alpha = 0.5, fill="blue" ) + 
        xlim( 0, 100 ) +
        xlab( "Concentration of Commercial Organizations" )


ggplot( dat.commercial.sub, aes( x=concentration ) ) + 
        geom_density( alpha = 0.5, fill="blue" ) + xlim( 0, 100 ) +
        xlab( "Concentration of Commercial Organizations" ) + facet_wrap( ~ NTMAJ12, nrow=3 ) + 
  theme_minimal()

ggplot( dat.commercial.sub, aes( y=concentration ) )  + 
        geom_boxplot( col="gray30", alpha=0.7) + 
        ylab( "Concentration of Commercial Organizations" ) +
        facet_wrap( ~ NTMAJ12, nrow=3 ) + theme_minimal()


```




## The Concentration of Commercial Nonprofits and Total Revenue


```{r}


jplot( log10(dat.commercial.sub$revenue), dat.commercial.sub$concentration, xlab="Total Revenue
       (logged)", ylab="The Concentration of Commercial Nonprofits", xaxt="n" )
axis( side=1, at=c(0,1,2,3,4,5,6,7,8,9,10,11),
      labels=c("1","10","100","1K","10K","100K","1M","10M","100M","1B","10B","100B") )




ggplot( dat.commercial.sub, aes(log10(revenue),concentration) )  + 
        geom_point( col="gray30", alpha=0.7) +
        xlab( "Total Revenue (logged)" ) + ylab( "The Concentration of Commercial Nonprofits" ) +
        theme_minimal()


ggplot( dat.commercial.sub, aes(log10(revenue),concentration) )  + 
        geom_point( col="gray30", alpha=0.7) +
        xlab( "Total Revenue (logged)" ) + ylab( "The Concentration of Commercial Nonprofits" ) +
        facet_wrap( ~ NTMAJ12, nrow=3 ) + theme_minimal() + geom_smooth(method='lm')


```


## The Density of Big Nonprofits and MSA Population


```{r}



```



# Reference











```{css, echo=F}
p {
color: black;
margin: 0 0 20px 0;
}

td {
    padding: 3px 10px 3px 10px;
    text-align: center;
}

table
{ 
    margin-left: auto;
    margin-right: auto;
    margin-top:80px;
    margin-bottom:100px;
}

h1, h2{
  margin-top:100px;
  margin-bottom:20px;
}

H5{
    text-align: center;
    color: gray;
    font-size:0.8em;
}

img {
    max-width: 90%;
    display: block;
    margin-right: auto;
    margin-left: auto;
    margin-top:30px;
    margin-bottom:20px;
}

pre {
  overflow-x: auto;
}

pre code {
   display: block; 
   padding: 0.5em;
   margin-bottom:20px;
}

code {
  font-size: 92%;
  border: 10px solid #F8F8F8;
  margin-bottom: 2px;
}

code[class] {
  background-color: #F8F8F8;
}

```

