---
title: "Markdown Template"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: haddock
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, warning=F, fig.width = 10)
```


# Load Packages

```{r}

library( haven )
library( tidyr )
library( dplyr )
library( ggplot2 )
library( ggthemes )
library( scales )
library( stargazer )
library( pander )


```

## Custom Functions

```{r}

jplot <- function( x1, x2, draw.line=T, ... )
{

	plot( x1, x2,
	      pch=19, 
	      col=gray(0.6, alpha = 0.4), 
	      cex=0.8,  
	      bty = "n",
	      cex.lab=1.5,
        ... )

	if( draw.line==T ){ 
		ok <- is.finite(x1) & is.finite(x2)
		lines( lowess(x2[ok]~x1[ok]), col="red", lwd=3 ) }

}

```



# CR4 Index



## Definitions

CR4 is defined as the summation of the market shares of the largest four firms. It is also common to see CR2 and CR8 (Seaman, Wilsker, Young, 2014).

## Calculations



CR4:

The summation of revenue of the largest four nonprofits devided by the summation of revenue of all nonprofits in a market.


CR2:

The summation of revenue of the largest two nonprofits devided by the summation of revenue of all nonprofits in a market

CR8:

The summation of revenue of the largest eight nonprofits devided by the summation of revenue of all nonprofits in a market


## Notes

We define a market as a Metropolitan Statistical Area (MSA) in which relevant nonprofits compete for resources to survive. The code to create the metrics is as follows.


# Create Metrics

```{r, eval=F}



# setwd("C:/Users/JK/Desktop/R Practice")
# core.2010 <- readRDS( "nccs.core2010pc.rds" )


core.2010 <- readRDS(gzcon(url("https://www.dropbox.com/s/byt7kh3kejg4ycl/nccs.core2010pc.rds?dl=1")))

names( core.2010 ) <- toupper( names( core.2010 ))

core <- select( core.2010, EIN, MSA_NECH, NTMAJ12, TOTREV, CONT)

core$NTMAJ12[ core$NTMAJ12 == "AR" ] <- "Arts"
core$NTMAJ12[ core$NTMAJ12 == "BH" ] <- "Universities"
core$NTMAJ12[ core$NTMAJ12 == "ED" ] <- "Education"
core$NTMAJ12[ core$NTMAJ12 == "EH" ] <- "Hospitals"
core$NTMAJ12[ core$NTMAJ12 == "EN" ] <- "Environmental"
core$NTMAJ12[ core$NTMAJ12 == "HE" ] <- "Health"
core$NTMAJ12[ core$NTMAJ12 == "HU" ] <- "Human Services"
core$NTMAJ12[ core$NTMAJ12 == "IN" ] <- "International"
core$NTMAJ12[ core$NTMAJ12 == "MU" ] <- "Mutual Benefit"
core$NTMAJ12[ core$NTMAJ12 == "PU" ] <- "Public Benefit"
core$NTMAJ12[ core$NTMAJ12 == "RE" ] <- "Religion"
core$NTMAJ12[ core$NTMAJ12 == "UN" ] <- "Unknown"

head(core) %>% pander()
             

core$NTMAJ12 <- factor( core$NTMAJ12 )





dat.CR4 <- 
  core %>% 
  group_by( MSA_NECH) %>%
  mutate(rank = rank(desc(TOTREV))) %>%
  arrange(rank) %>%
  summarize( n=n(), revenue= sum(TOTREV), top1 = sum(nth(TOTREV, 1)), top2 = sum(nth(TOTREV, 2)), 
             top3 = sum(nth(TOTREV, 3)), top4 = sum(nth(TOTREV, 4)), 
             percent=(top1+top2+top3+top4)/revenue)




dat.CR4.msa.sub <- 
  core %>% 
  group_by( MSA_NECH, NTMAJ12) %>% 
  mutate(rank = rank(desc(TOTREV))) %>%
  arrange(rank) %>%
  summarize( n=n(), revenue= sum(TOTREV), top1 = sum(nth(TOTREV, 1)), top2 = sum(nth(TOTREV, 2)), 
             top3 = sum(nth(TOTREV, 3)), top4 = sum(nth(TOTREV, 4)), 
             percent=(top1+top2+top3+top4)/revenue)


dat.CR2 <- 
  core %>% 
  group_by( MSA_NECH) %>% 
  mutate(rank = rank(desc(TOTREV))) %>%
  arrange(rank) %>%
  summarize( n=n(), revenue= sum(TOTREV), top1 = sum(nth(TOTREV, 1)), top2 = sum(nth(TOTREV, 2)), 
             percent=(top1+top2)/revenue)


dat.CR2.msa.sub <- 
  core %>% 
  group_by( MSA_NECH, NTMAJ12) %>% 
  mutate(rank = rank(desc(TOTREV))) %>%
  arrange(rank) %>%
  summarize( n=n(), revenue= sum(TOTREV), top1 = sum(nth(TOTREV, 1)), top2 = sum(nth(TOTREV, 2)), 
             percent=(top1+top2)/revenue)


dat.CR8 <- 
  core %>% 
  group_by( MSA_NECH) %>% 
  mutate(rank = rank(desc(TOTREV))) %>%
  arrange(rank) %>%
  summarize( n=n(), revenue= sum(TOTREV), top1 = sum(nth(TOTREV, 1)), top2 = sum(nth(TOTREV, 2)),    
             top3 = sum(nth(TOTREV, 3)), top4 = sum(nth(TOTREV, 4)), top5 = sum(nth(TOTREV, 5)), 
             top6 = sum(nth(TOTREV, 6)), top7 = sum(nth(TOTREV, 7)), top8 = sum(nth(TOTREV, 8)), 
             percent=(top1+top2+top3+top4+top5+top6+top7+top8)/revenue)
                                                                                                                                                                                             
dat.CR8.msa.sub <- 
  core %>% 
  group_by( MSA_NECH, NTMAJ12) %>%
  mutate(rank = rank(desc(TOTREV))) %>%
  arrange(rank) %>%
  summarize( n=n(), revenue= sum(TOTREV), top1 = sum(nth(TOTREV, 1)), top2 = sum(nth(TOTREV, 2)),    
             top3 = sum(nth(TOTREV, 3)), top4 = sum(nth(TOTREV, 4)), top5 = sum(nth(TOTREV, 5)), 
             top6 = sum(nth(TOTREV, 6)), top7 = sum(nth(TOTREV, 7)), top8 = sum(nth(TOTREV, 8)), 
             percent=(top1+top2+top3+top4+top5+top6+top7+top8)/revenue)

```

# Descriptive Statistics

## HHI by MSAs

```{r}

####Something wrong here, too many NAs

dat.CR4 <- 
  core %>% 
  group_by( MSA_NECH) %>% 
  mutate(rank = rank(desc(TOTREV))) %>%
  arrange(rank) %>%
  summarize( n=n(), revenue= sum(TOTREV), top1 = sum(nth(TOTREV, 1)), top2 = sum(nth(TOTREV, 2)), 
             top3 = sum(nth(TOTREV, 3)), top4 = sum(nth(TOTREV, 4)), 
             percent=(top1+top2+top3+top4)/revenue)

core$TOTREV[ core$TOTREV < 0 ] <- 0

summary(dat.CR4)


ggplot( dat.CR4, aes(x = percent )) +  geom_density( alpha = 0.5, fill="blue" ) + xlim( 0, 1 ) 



```

## HHI by Subsector

```{r}


####Something wrong here, too many NAs

dat.sub <- 
  core %>% 
  group_by( NTMAJ12) %>% 
  mutate(rank = rank(desc(TOTREV))) %>%
  arrange(rank) %>%
  summarize( n=n(), revenue= sum(TOTREV), top1 = sum(nth(TOTREV, 1)), top2 = sum(nth(TOTREV, 2)), 
             top3 = sum(nth(TOTREV, 3)), top4 = sum(nth(TOTREV, 4)), 
             percent=(top1+top2+top3+top4)/revenue)



summary(dat.sub)

ggplot( dat.CR4.msa.sub, aes(x = percent)) +  geom_density( alpha = 0.5, fill="blue" ) + xlim( -0.05, 1 ) +xlab( "CR4" )


ggplot( dat.CR4.msa.sub, aes( x= percent ) ) + 
        geom_density( alpha = 0.5, fill="blue" ) + xlim( -0.05, 1 ) +
        xlab( "CR4" ) + facet_wrap( ~ NTMAJ12, nrow=3 ) + theme_minimal()

ggplot( dat.CR4.msa.sub, aes(y = percent) )  + 
        geom_boxplot( col="gray30", alpha=0.7) + 
        ylab( "CR4 Across MSAs" ) +
        facet_wrap( ~ NTMAJ12, nrow=3 ) + theme_minimal()


```



## CR4 and Subsector Size

```{r}


jplot( log10(dat.CR4.msa.sub$n), dat.CR4.msa.sub$percent, xlab="Number of Nonprofits", ylab="CR4", xaxt="n" )
axis( side=1, at=c(0,1,2,3,4), labels=c("1","10","100","1,000","10,000") )



ggplot( dat.CR4.msa.sub, aes(log10(n),percent) )  + 
        geom_point( col="gray30", alpha=0.7) + 
        xlab( "Number of Nonprofits (logged)" ) + ylab( "CR4 Across MSAs" ) +
        theme_minimal()


ggplot( dat.CR4.msa.sub, aes(log10(n),percent) )  + 
        geom_point( col="gray30", alpha=0.7) + 
        xlab( "Number of Nonprofits (logged)" ) + ylab( "CR4 Across MSAs" ) +
        facet_wrap( ~ NTMAJ12, nrow=3 ) + theme_minimal()





```





# Reference



Seaman, B., A. Wilsker, and D. R. Young. 2014. ¡¥¡¥Measuring Concentration and Competition in the U.S. Nonprofit Sector: Implications for Research and Public Policy.¡¦¡¦ Nonprofit Policy Forum 5(2): 231¡V259.

Castaneda, M. A., Garen, J., & Thornton, J. (2007). Competition, contractibility, and the market for donors to nonprofits. The Journal of Law, Economics, & Organization, 24(1), 215-246.


Feigenbaum, S. (1987). Competition and performance in the nonprofit sector: The case of US medical research charities. The Journal of Industrial Economics, 241-253.






```{css, echo=F}
p {
color: black;
margin: 0 0 20px 0;
}

td {
    padding: 3px 10px 3px 10px;
    text-align: center;
}

table
{ 
    margin-left: auto;
    margin-right: auto;
    margin-top:80px;
    margin-bottom:100px;
}

h1, h2{
  margin-top:100px;
  margin-bottom:20px;
}

H5{
    text-align: center;
    color: gray;
    font-size:0.8em;
}

img {
    max-width: 90%;
    display: block;
    margin-right: auto;
    margin-left: auto;
    margin-top:30px;
    margin-bottom:20px;
}

pre {
  overflow-x: auto;
}

pre code {
   display: block; 
   padding: 0.5em;
   margin-bottom:20px;
}

code {
  font-size: 92%;
  border: 10px solid #F8F8F8;
  margin-bottom: 2px;
}

code[class] {
  background-color: #F8F8F8;
}

```

