---
title: "Markdown Template"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: haddock
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, warning=F, fig.width = 10)
```


# Load Packages

```{r}

library( haven )
library( tidyr )
library( dplyr )
library( ggplot2 )
library( ggthemes )
library( scales )
library( stargazer )
library( pander )


```

## Custom Functions

```{r}

jplot <- function( x1, x2, draw.line=T, ... )
{

	plot( x1, x2,
	      pch=19, 
	      col=gray(0.6, alpha = 0.4), 
	      cex=0.8,  
	      bty = "n",
	      cex.lab=1.5,
        ... )

	if( draw.line==T ){ 
		ok <- is.finite(x1) & is.finite(x2)
		lines( lowess(x2[ok]~x1[ok]), col="red", lwd=3 ) }

}

```



# Kwoka Index



## Definitions



## Calculations

HHI can be calculated as follows: 

Approach 1:



Approach 2:



Approach 3:



## Notes

We define a market as a Metropolitan Statistical Area (MSA) in which relevant nonprofits compete for resources to survive. The code to create HHI metrics is as follows.


# Create Metrics

```{r, eval=F}



setwd("C:/Users/JK/Desktop/R Practice")

core.2010 <- readRDS( "nccs.core2010pc.rds" )

names( core.2010 ) <- toupper( names( core.2010 ))

core <- select( core.2010, EIN, MSA_NECH, NTMAJ12, TOTREV, CONT)

core$NTMAJ12[ core$NTMAJ12 == "AR" ] <- "Arts"
core$NTMAJ12[ core$NTMAJ12 == "BH" ] <- "Universities"
core$NTMAJ12[ core$NTMAJ12 == "ED" ] <- "Education"
core$NTMAJ12[ core$NTMAJ12 == "EH" ] <- "Hospitals"
core$NTMAJ12[ core$NTMAJ12 == "EN" ] <- "Environmental"
core$NTMAJ12[ core$NTMAJ12 == "HE" ] <- "Health"
core$NTMAJ12[ core$NTMAJ12 == "HU" ] <- "Human Services"
core$NTMAJ12[ core$NTMAJ12 == "IN" ] <- "International"
core$NTMAJ12[ core$NTMAJ12 == "MU" ] <- "Mutual Benefit"
core$NTMAJ12[ core$NTMAJ12 == "PU" ] <- "Public Benefit"
core$NTMAJ12[ core$NTMAJ12 == "RE" ] <- "Religion"
core$NTMAJ12[ core$NTMAJ12 == "UN" ] <- "Unknown"

head(core) %>% pander()
             

core$NTMAJ12 <- factor( core$NTMAJ12 )

dat.kindex3 <- 
  core %>% 
  group_by( MSA_NECH) %>%
  mutate(rank = rank(desc(TOTREV))) %>%
  arrange(rank) %>%
  summarize( n=n(), top1 = sum(nth(TOTREV, 1)), top2 = sum(nth(TOTREV, 2)), 
             top3 = sum(nth(TOTREV, 3)),total = top1 + top2 + top3, Kindex3 = (((top1/total)^2 +
                                                                                 (top2/total)^2+
                                                                               (top3/total)^2)))
View(dat.kindex3)



dat.kindex2 <- 
  core %>% 
  group_by( MSA_NECH) %>%
  mutate(rank = rank(desc(TOTREV))) %>%
  arrange(rank) %>%
  summarize( n=n(), top1 = sum(nth(TOTREV, 1)), top2 = sum(nth(TOTREV, 2)), total = top1 + top2, 
             Kindex2 = (((top1/total)^2 + (top2/total)^2)))




dat.kindex3.sub <- 
  core %>% 
  group_by( MSA_NECH, NTMAJ12) %>%
  mutate(rank = rank(desc(TOTREV))) %>%
  arrange(rank) %>%
  summarize( n=n(), revenue = sum(TOTREV), top1 = sum(nth(TOTREV, 1)), top2 = sum(nth(TOTREV, 2)), 
             top3 = sum(nth(TOTREV, 3)),total = top1 + top2 + top3, Kindex3 = (((top1/total)^2 +
                                                                                 (top2/total)^2+
                                                                               (top3/total)^2)))

View(dat.kindex3.sub)


dat.kindex3.sub$revenue[ dat.kindex3.sub$revenue < 0 ] <- 0

summary(dat.kindex3.sub$Kindex3)
summary(core$TOTREV)




```

# Descriptive Statistics

## Kwoka Index (3) by MSAs

```{r}

dat.kindex3 <- 
  core %>% 
  group_by( MSA_NECH) %>%
  mutate(rank = rank(desc(TOTREV))) %>%
  arrange(rank) %>%
  summarize( n=n(), top1 = sum(nth(TOTREV, 1)), top2 = sum(nth(TOTREV, 2)), 
             top3 = sum(nth(TOTREV, 3)),total = top1 + top2 + top3, KwokaIndex = (((top1/total)^2 +
                                                                                 (top2/total)^2+
                                                                               (top3/total)^2)))




ggplot( dat.kindex3, aes(x = KwokaIndex )) +  geom_density( alpha = 0.5, fill="blue" ) + xlim( 0, 1 ) 



```

## HHI by Subsector

```{r}

dat.kindex3.sub <- 
  core %>% 
  group_by(NTMAJ12) %>%
  mutate(rank = rank(desc(TOTREV))) %>%
  arrange(rank) %>%
  summarize( n=n(), top1 = sum(nth(TOTREV, 1)), top2 = sum(nth(TOTREV, 2)), 
             top3 = sum(nth(TOTREV, 3)),total = top1 + top2 + top3, KwokaIndex = (((top1/total)^2 +
                                                                                 (top2/total)^2+
                                                                               (top3/total)^2)))


dat.kindex3.sub <- 
  core %>% 
  group_by( MSA_NECH, NTMAJ12) %>%
  mutate(rank = rank(desc(TOTREV))) %>%
  arrange(rank) %>%
  summarize( n=n(), revenue = sum(TOTREV), contribution = sum(CONT), top1 = sum(nth(TOTREV, 1)), top2 =
               sum(nth(TOTREV, 2)), top3 = sum(nth(TOTREV, 3)),total = top1 + top2 + top3, KwokaIndex =
               (((top1/total)^2 + (top2/total)^2+ (top3/total)^2)))

ggplot( dat.kindex3.sub, aes(x = KwokaIndex)) +  geom_density( alpha = 0.5, fill="blue" ) + xlim( 
  -0.05, 1 ) + xlab( "Kwoka Index" )


ggplot( dat.kindex3.sub, aes( x=KwokaIndex ) ) + 
        geom_density( alpha = 0.5, fill="blue" ) + xlim( -0.05, 1 ) +
        xlab( "Kwoka Index" ) + facet_wrap( ~ NTMAJ12, nrow=3 ) + theme_minimal()

ggplot( dat.kindex3.sub, aes(y = KwokaIndex) )  + 
        geom_boxplot( col="gray30", alpha=0.7) + 
        ylab( "Kwoka Index Across MSAs" ) +
        facet_wrap( ~ NTMAJ12, nrow=3 ) + theme_minimal()


```



## Kwoka Index and Subsector Size

```{r}


jplot( log10(dat.kindex3.sub$n), dat.kindex3.sub$KwokaIndex, xlab="Number of Nonprofits", ylab="Kwoka Index", xaxt="n") 
axis( side=1, at=c(0,1,2,3,4), labels=c("1","10","100","1,000","10,000") )



ggplot( dat.kindex3.sub, aes(log10(n),KwokaIndex) )  + 
        geom_point( col="gray30", alpha=0.7) + 
        xlab( "Number of Nonprofits (logged)" ) + ylab( "Kwoka Index Across MSAs" ) +
        theme_minimal()


ggplot( dat.kindex3.sub, aes(log10(n),KwokaIndex) )  + 
        geom_point( col="gray30", alpha=0.7) + 
        xlab( "Number of Nonprofits (logged)" ) + ylab( "Kwoka Index Across MSAs" ) +
        facet_wrap( ~ NTMAJ12, nrow=3 ) + theme_minimal()





```

## Kwoka Index and Total Revenue


```{r}


jplot( log10(dat.kindex3.sub$revenue), dat.kindex3.sub$KwokaIndex, xlab="Total Revenue (logged)", 
       ylab="Kwoka Index", xaxt="n" )
axis( side=1, at=c(0,1,2,3,4,5,6,7,8,9,10,11),
      labels=c("1","10","100","1K","10K","100K","1M","10M","100M","1B","10B","100B") )




ggplot( dat.kindex3.sub, aes(log10(revenue),KwokaIndex) )  + 
        geom_point( col="gray30", alpha=0.7) +
        xlab( "Total Revenue (logged)" ) + ylab( "Kwoka Index Across MSAs" ) +
        theme_minimal()


ggplot( dat.kindex3.sub, aes(log10(revenue),KwokaIndex) )  + 
        geom_point( col="gray30", alpha=0.7) +
        xlab( "Total Revenue (logged)" ) + ylab( "Kwoka Index Across MSAs" ) +
        facet_wrap( ~ NTMAJ12, nrow=3 ) + theme_minimal()



ggplot( dat.kindex3.sub, aes(log10(revenue),KwokaIndex) )  + 
        geom_point( col="gray30", alpha=0.7) +
        geom_smooth(method="lm") +
        xlab( "Total Revenue (logged)" ) + ylab( "Kwoka Index Across MSAs" ) +
        facet_wrap( ~ NTMAJ12, nrow=3 ) + theme_minimal()

```


## Kwoka Index and Subsector Contribution


```{r}

jplot( log10(dat.kindex3.sub$contribution), dat.kindex3.sub$KwokaIndex, xlab="Contribution (logged)", 
       ylab="Kwoka Index", xaxt="n" )
axis( side=1, at=c(0,1,2,3,4,5,6,7,8,9,10,11), 
      labels=c("1","10","100","1K","10K","100K","1M","10M","100M","1B","10B","100B") )


ggplot( dat.kindex3.sub, aes(log10(contribution),KwokaIndex) )  + 
        geom_point( col="gray30", alpha=0.7) +
        xlab( "Contrubution (logged)" ) + ylab( "Kwoka Index Across MSAs" ) +
        theme_minimal()


ggplot( dat.kindex3.sub, aes(log10(contribution),KwokaIndex) )  + 
        geom_point( col="gray30", alpha=0.7) +
        xlab( "Contrubution (logged)" ) + ylab( "Kwoka Index Across MSAs" ) +
        facet_wrap( ~ NTMAJ12, nrow=3 ) + theme_minimal()


ggplot( dat.kindex3.sub, aes(log10(contribution),KwokaIndex) )  + 
        geom_point( col="gray30", alpha=0.7) +
        geom_smooth(method="lm") +
        xlab( "Contrubution (logged)" ) + ylab( "Kwoka Index Across MSAs" ) +
        facet_wrap( ~ NTMAJ12, nrow=3 ) + theme_minimal()
```



# Reference



Seaman, B., A. Wilsker, and D. R. Young. 2014. ¡¥¡¥Measuring Concentration and Competition in the U.S. Nonprofit Sector: Implications for Research and Public Policy.¡¦¡¦ Nonprofit Policy Forum 5(2): 231¡V259.









```{css, echo=F}
p {
color: black;
margin: 0 0 20px 0;
}

td {
    padding: 3px 10px 3px 10px;
    text-align: center;
}

table
{ 
    margin-left: auto;
    margin-right: auto;
    margin-top:80px;
    margin-bottom:100px;
}

h1, h2{
  margin-top:100px;
  margin-bottom:20px;
}

H5{
    text-align: center;
    color: gray;
    font-size:0.8em;
}

img {
    max-width: 90%;
    display: block;
    margin-right: auto;
    margin-left: auto;
    margin-top:30px;
    margin-bottom:20px;
}

pre {
  overflow-x: auto;
}

pre code {
   display: block; 
   padding: 0.5em;
   margin-bottom:20px;
}

code {
  font-size: 92%;
  border: 10px solid #F8F8F8;
  margin-bottom: 2px;
}

code[class] {
  background-color: #F8F8F8;
}

```

