---
title: "Markdown Template"
output:
  html_document:
    code_folding: show
    df_print: paged
    highlight: haddock
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=F, warning=F, fig.width = 10)
```


# Load Packages

```{r}

library( haven )
library( tidyr )
library( dplyr )
library( ggplot2 )
library( ggthemes )
library( scales )
library( stargazer )
library( pander )


```

## Custom Functions

```{r}

jplot <- function( x1, x2, draw.line=T, ... )
{

	plot( x1, x2,
	      pch=19, 
	      col=gray(0.6, alpha = 0.4), 
	      cex=0.8,  
	      bty = "n",
	      cex.lab=1.5,
        ... )

	if( draw.line==T ){ 
		ok <- is.finite(x1) & is.finite(x2)
		lines( lowess(x2[ok]~x1[ok]), col="red", lwd=3 ) }

}

```



# The Concentration of Small Nonprofit Index



## Definitions



## Calculations



Approach 1:



Approach 2:



Approach 3:



## Notes

We define a market as a Metropolitan Statistical Area (MSA) in which relevant nonprofits compete for resources to survive. The code to create the metrics is as follows.


# Create Metrics

```{r, eval=F}



# setwd("C:/Users/JK/Desktop/R Practice")
# core.2010 <- readRDS( "nccs.core2010pc.rds" )

core.2010 <- readRDS(gzcon(url("https://www.dropbox.com/s/byt7kh3kejg4ycl/nccs.core2010pc.rds?dl=1")))


names( core.2010 ) <- toupper( names( core.2010 ))

core <- select( core.2010, EIN, MSA_NECH, NTMAJ12, TOTREV, CONT)

core$NTMAJ12[ core$NTMAJ12 == "AR" ] <- "Arts"
core$NTMAJ12[ core$NTMAJ12 == "BH" ] <- "Universities"
core$NTMAJ12[ core$NTMAJ12 == "ED" ] <- "Education"
core$NTMAJ12[ core$NTMAJ12 == "EH" ] <- "Hospitals"
core$NTMAJ12[ core$NTMAJ12 == "EN" ] <- "Environmental"
core$NTMAJ12[ core$NTMAJ12 == "HE" ] <- "Health"
core$NTMAJ12[ core$NTMAJ12 == "HU" ] <- "Human Services"
core$NTMAJ12[ core$NTMAJ12 == "IN" ] <- "International"
core$NTMAJ12[ core$NTMAJ12 == "MU" ] <- "Mutual Benefit"
core$NTMAJ12[ core$NTMAJ12 == "PU" ] <- "Public Benefit"
core$NTMAJ12[ core$NTMAJ12 == "RE" ] <- "Religion"
core$NTMAJ12[ core$NTMAJ12 == "UN" ] <- "Unknown"

head(core) %>% pander()
             

core$NTMAJ12 <- factor( core$NTMAJ12 )

####"percentage with revenue less than $50k in a market" is a great idea. The problem is that there are MSAs only having one nonprofit and the nonprofit's revenue is less than $50k. Its percentage would be 100%. It looks like the nonprofit is located in a competative MSA (100%); however, the nonprofit may be located in a suburban area where not too many people need its products and services. 

dat.con.small.rev50k <- 
  core %>% 
  group_by( MSA_NECH) %>% 
  summarize(n=n(), n_small=sum(TOTREV < 50000), percent=n_small/n*100 )

dat.con.small.rev50k %>%  pander()

hist( dat.con.small.rev50k$percent, breaks=100, col="gray", border="white" )
 

dat.con.small.rev100k <- 
  core %>% 
  group_by( MSA_NECH) %>% 
  summarize(n=n(), n_small=sum(TOTREV < 100000), percent=n_small/n*100 )

hist( dat.con.small.rev100k$percent, breaks=100, col="gray", border="white" )
  

plot( dat.con.small.rev50k$percent, dat.con.small.rev100k$percent, pch=19 )



dat.con.small.rev50k.msa.sub <- 
  core %>% 
  group_by( MSA_NECH, NTMAJ12 ) %>% 
  summarize(n=n(), n_small=sum(TOTREV < 50000), percent=n_small/n*100 ) 

dat.con.small.rev50k.msa.sub  %>% pander()

hist( dat.con.small.rev50k.msa.sub$percent, breaks=100, col="gray", border="white" )
 

dat.con.small.rev50k.sub <- 
  core %>% 
  group_by( NTMAJ12 ) %>% 
  summarize(n=n(), n_small=sum(TOTREV < 50000), percent=n_small/n*100 ) 

dat.con.small.rev50k.sub %>% pander()


###So,what about we ues "percentage with revenue less than the average revenue in a market"

dat.con.small.revmean <- 
  core %>% 
  group_by( MSA_NECH ) %>% 
  summarize( ave.rev=mean(TOTREV),n=n(), n_small=sum(TOTREV < mean(TOTREV)), percent=n_small/n*100)%>%
  pander()


dat.con.small.revmean.msa.sub <- 
  core %>% 
  group_by( MSA_NECH, NTMAJ12 ) %>% 
  summarize( ave.rev=mean(TOTREV),n=n(), n_small=sum(TOTREV < mean(TOTREV)), percent=n_small/n*100, revenue=sum(TOTREV))


```

# Descriptive Statistics

## small nonprofits by MSAs and MSAs/Sub

```{r}

dat.msa <-
  core %>%
  group_by(MSA_NECH) %>%
  summarize( ave.rev=mean(TOTREV),n=n(), n_small=sum(TOTREV < mean(TOTREV)), percent=n_small/n*100 ) 

ggplot( dat.msa, aes(x=percent)) +  geom_histogram()

ggplot(dat.msa, aes(x=percent)) +  geom_density( alpha = 0.5, fill="blue" ) + xlim( 0, 100 )


dat.msa.sub <-
  core %>%
  group_by(MSA_NECH,NTMAJ12) %>%
  summarize( ave.rev=mean(TOTREV),n=n(), n_small=sum(TOTREV < mean(TOTREV)), percent=n_small/n*100 )

ggplot( dat.msa.sub, aes(x=percent)) +  geom_histogram()

ggplot(dat.msa.sub, aes(x=percent)) +  geom_density( alpha = 0.5, fill="blue" ) + xlim( 0, 100 )


```

## small organization by Subsector

```{r}

dat.sub <-
core %>%
group_by(NTMAJ12) %>%
summarize( ave.rev=mean(TOTREV),n=n(), n_small=sum(TOTREV < mean(TOTREV)), percent=n_small/n*100 )

summary(dat.sub)

ggplot( dat.con.small.revmean.msa.sub, aes(x = percent)) +  
        geom_density( alpha = 0.5, fill="blue" ) + 
        xlim( 0, 100 ) +
        xlab( "Percentage of Small Organizations" )


ggplot( dat.con.small.revmean.msa.sub, aes( x=percent ) ) + 
        geom_density( alpha = 0.5, fill="blue" ) + xlim( 0, 100 ) +
        xlab( "Percentage of Small Organizations" ) + facet_wrap( ~ NTMAJ12, nrow=3 ) + theme_minimal()

ggplot( dat.con.small.revmean.msa.sub, aes( y=percent ) )  + 
        geom_boxplot( col="gray30", alpha=0.7) + 
        ylab( "Percentage of Small Organizations" ) +
        facet_wrap( ~ NTMAJ12, nrow=3 ) + theme_minimal()


```




## The Concentration of Small Nonprofits and Total Revenue


```{r}


jplot( log10(dat.con.small.revmean.msa.sub$revenue), dat.con.small.revmean.msa.sub$percent, xlab="Total Revenue (logged)", ylab="The Concentration of Small Nonprofits", xaxt="n" )
axis( side=1, at=c(0,1,2,3,4,5,6,7,8,9,10,11), labels=c("1","10","100","1K","10K","100K","1M","10M","100M","1B","10B","100B") )




ggplot( dat.con.small.revmean.msa.sub, aes(log10(revenue),percent) )  + 
        geom_point( col="gray30", alpha=0.7) +
        xlab( "Total Revenue (logged)" ) + ylab( "The Concentration of Small Nonprofits" ) +
        theme_minimal()


ggplot( dat.con.small.revmean.msa.sub, aes(log10(revenue),percent) )  + 
        geom_point( col="gray30", alpha=0.7) +
        xlab( "Total Revenue (logged)" ) + ylab( "The Concentration of Small Nonprofits" ) +
        facet_wrap( ~ NTMAJ12, nrow=3 ) + theme_minimal()

```


## The Concentration of Small Nonprofits and MSA Population


```{r}



```



# Reference











```{css, echo=F}
p {
color: black;
margin: 0 0 20px 0;
}

td {
    padding: 3px 10px 3px 10px;
    text-align: center;
}

table
{ 
    margin-left: auto;
    margin-right: auto;
    margin-top:80px;
    margin-bottom:100px;
}

h1, h2{
  margin-top:100px;
  margin-bottom:20px;
}

H5{
    text-align: center;
    color: gray;
    font-size:0.8em;
}

img {
    max-width: 90%;
    display: block;
    margin-right: auto;
    margin-left: auto;
    margin-top:30px;
    margin-bottom:20px;
}

pre {
  overflow-x: auto;
}

pre code {
   display: block; 
   padding: 0.5em;
   margin-bottom:20px;
}

code {
  font-size: 92%;
  border: 10px solid #F8F8F8;
  margin-bottom: 2px;
}

code[class] {
  background-color: #F8F8F8;
}

```

